\LARGE \bf Cascading Situation Assessment for Robots

### 2017-02-27 paper.tex
WIP on intro, in particular, related work

\uwds is a distributed and lightweight framework that enables robot programmers
to build and refine in real-time spatial and temporal models of the environment
surrounding a robot. \uwds makes it easy, then, to share these models amongst
the other software components running on the robot, to eventually provide a
distributed environment model for the robot.

The components which require such spatial and temporal models of the environment
are typically found in the intermediate layers of robotic architectures, between
the low-level perceptual layers, and the high-level decisional layers. They
typically include geometric reasoners (that compute spatial and topological
relations between objects), motion planners, gestures/sequence recognizers.

Difficulties however arise quickly. These different components might need
different models, at different spatial and temporal resolutions. A 3D motion
planner would typically use coarse 3D models of surrounding objects to lower the
computational load while planning, while a spatial reasoner might need on the
contrary a high-resolution model of the environment for accurate rendering when
attempting to estimate the visibility of an object. Spatial models are also
usually large as they might contain a lot of geometric data (including point
clouds and meshes). Efficient sharing of such datasets is non-trivial,
particularly when the said data are dynamically acquired (\eg in the case of
map building or object recognition).

Traditional robotic middlewares like ROS are not particularly well-suited to
deal with these complexities: geometric data can be represented, but are not
first-class citizens (basic tasks like displaying a 3D mesh at a dynamic
position are non-trivial with ROS, for instance) and representing and reasoning
on \emph{alternative states} of the environment is not directly doable.

@@

\subsection{Cascading Situation Assessment}

Anchoring perceptions in a symbolic model requires perception abilities and
their symbolic interpretation. We call \emph{physical situation assessment} the
cognitive skill that a robot exhibits when it assesses the nature and content of its
surroundings and monitors its evolution.

@@

\section{Related work}

\cite{sisbot2011situation}
\cite{naef2003blue}
\cite{bustos2016unified}

\uwds is inspired by geometric and temporal reasoning components like {\sc
Spark} (\emph{SPAtial Reasoning \&
Knowledge}~\cite{sisbot2011situation} (see~\cite{lemaignan2016artificial} for a
high-level description of the integration of a component like {\sc Spark} in a
complete robotic architecture).

It acts as a situation assessment reasoner that generates symbolic knowledge from the
geometry of the environment with respect to relations between objects, robots
and humans, also
taking into account the different perspective that each agent has on the
environment.  {\sc Spark} embeds an \emph{amodal} (as defined by Mavridis and
Roy in~\cite{Mavridis2006}: the different perceptual modalities are abstracted
away into a blended spatial model) geometric model of the environment that
serves both as basis for the fusion of the perception modalities and as bridge
with the symbolic layer. This geometric model is built from 3D CAD models of the
objects, furnitures and robots, and full body, rigged models of humans
(Figure~\ref{fig:sparkScreenshot}).  It is updated at run-time by the robot's
sensors (usually, a combination of vision-based tracking of 2D fiducial markers
to identify and localise objects, and Kinect-based skeleton tracking of humans,
optionally assisted by motion capture to accurately track the head motion, which
is required to compute what the human is looking at).

\subsection*{Relation to existing robotic middlewares}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{overview}
    \caption{Schema of a possible \uwds network: eight possible, user-written
    \emph{clients} (in blue) are exchanging environment models through four shared
    \emph{worlds} (made of joint spatial and temporal models). This architecture
    enables successive and modular refinement of the models
    (\emph{cascading} situation assessment), effectively adapted to each
    client's needs (yellow nodes represent other possible components in the system that do not
    interact directly with the \uwds network).}

>>>>

\uwds is a distributed and lightweight framework that enables robot programmers
to build and refine in real-time spatial and temporal models of the environment
surrounding a robot. \uwds make it possible to share these world models amongst
the software components running on the robot. Critically, \uwds permits to
represent and manipulate in a distributed manner \emph{multiple alternatives} to
the current, perceived world model: for instance, the world with some objects filtered out,
the world `viewed' from the perspective of another agent, an hypothetical world
resulting from the simulated application of a plan, etc.

\subsection{Situation Assessment}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{overview}
    \caption{Schema of a possible \uwds network: eight possible, user-written
    \emph{clients} (in blue) are exchanging environment models through four shared
    \emph{worlds} (made of joint spatial and temporal models). This architecture
    enables successive and modular refinement of the models
    (\emph{cascading} situation assessment), effectively adapted to each
    client's needs (yellow nodes represent other possible components in the system that do not
    interact directly with the \uwds network).}

Anchoring perceptions in a symbolic model suitable for decision-making requires
perception abilities and their symbolic interpretation. We call \emph{physical
situation assessment} the cognitive skill that a robot exhibits when it
represents and assesses the nature and content of its surroundings and monitors
its evolution.

\uwds is specifically inspired by the geometric and temporal reasoning module
{\sc Spark} (\emph{SPAtial Reasoning \& Knowledge}~\cite{sisbot2011situation}.
{\sc Spark} acts as a situation assessment reasoner that generates symbolic
knowledge from the geometry of the environment with respect to relations between
objects, robots and humans. It also takes into account the different perspective
that each agent has on the environment. {\sc Spark} embeds a
modality-independent geometric model of the environment that serves both as
basis for the fusion of the perception modalities and as bridge with the
symbolic layer~\cite{lemaignan2016artificial}. This geometric model is built
from 3D CAD models of the objects, furnitures and robots, and full body, rigged
models of humans.  It is updated at run-time by the robot's sensors.
Likewise, \uwds embeds a grounded amodal model of the environment, updated
online from the robot's sensors (sensor fusion).

However, {\sc Spark} is a monolithic module, that does not aim at sharing its
internal 3D model to other external components. On the contrary, \uwds focuses
on offering a shared and distributed representation within the robot's software
architecture. Previous work on distributed scene graphs~\cite{naef2003blue} has
been applied to robotics to provide a shared 3D representation of the robot's
environment (like the \emph{Robot Scene Graph}~\cite{blumenthal2013scene} or 
\emph{Deep State Representation}~\cite{bustos2016unified}).

\uwds further extends this line of work by providing the ability to manipulate
several \emph{alternative 3D worlds}, corresponding to filtered or hypothetical
\emph{views} on the perceived model of the environment.

\subsection{Representing Alternative States of the Environment}

The components which make use of spatial and temporal models of the environment
are usually found in the intermediate layers of robotic architectures, between
the low-level perceptual layers, and the high-level decisional layers. They
include modules like geometric reasoners (that compute spatial and topological
relations between objects), motion planners, action recognition.

These different components typically need different models, at different spatial
and temporal resolutions. A 3D motion planner would typically use coarse 3D
models of surrounding objects to lower the computational load while planning,
while a spatial reasoner might need on the contrary a high-resolution model of
the environment for accurate rendering when attempting to estimate the
visibility of an object. Spatial models are also usually large as they might
contain a lot of geometric data (including point clouds and meshes). Efficient
sharing of such datasets is non-trivial, particularly when the said data are
dynamically acquired (\eg in the case of map building or object recognition).

Traditional robotic middlewares like ROS are not particularly well-suited to
deal with these complexities: geometric data can be represented, but are not
first-class citizens (basic tasks like displaying a 3D mesh at a dynamic
position are non-trivial with ROS, for instance) and representing and reasoning
on \emph{alternative states} of the environment is not directly feasible.

@@

@@

Relation to Beetz's \emph{deep representation}~\cite{beetz2010towards}

GSM (for \emph{Grounded Situation Model})~\cite{Mavridis2006} is a knowledge
representation system primarily built to ``facilitate cross-modal
interoperability'',  especially in the context of verbal interaction with a
robot.

GSM does not rely on any formal language but rather on a layered data structure
(figure~\ref{fig|gsm}) that organises the surrounding world into agents and
relations between agents.  Each agent (any animate or inanimate object) is
attached to a physical model (made of \emph{body parts} that have properties
like their position, color, etc.) and a mental model (which is a recursively
embedded GSM, thus allowing a sort of theory of mind).

Properties are represented in three layers: a stochastic representation, close
to sensory percepts, a \emph{continuous single-valued} encoding of the
stochastic model, and a discrete, categorical model.

One notable feature of GSM is the \emph{bidirectionality} of the grounding
process: not only sensor percepts are abstracted into categories suitable for
human conversation, but human utterance (like ``There is a red ball in the
center of the table'') can also be turned into property descriptions. This
basically enable the knowledge representation system of the robot to
\emph{imagine} entities.

GSM also features several strategies for managing time and events.
\emph{Moments} are created by storing timestamped snap-shots of GSM, and
\emph{event classifiers} allow to define and detect events.

\subsection{Relation to existing robotic middlewares}

@@
